---
title: "Final Individual Report"
author: "Giulio Orazzo"
date: last-modified
project:
  type: website
  output-dir: docs
format :
  html:
    code-fold: true
    code-summary: "Show the code"
    toc: true
    toc-location: right
    toc-title: "On this page"
    css: styles.css
    html-math-method: mathjax
execute:
  echo: true  
  warning: false
  message: false

draft: false

---
## Table of Contents

1. [Introduction](#introduction)
2. 
[Assets Analysis](#assets-analysis)


# Introduction{#introduction}

## Research Question

#### Are developed international markets more effective than emerging markets for long-term portfolio growth, and how do they compare to the United States?

This project investigates whether developed international equity markets (e.g., Europe, Japan) outperform emerging markets (e.g., China, India, Brazil) in terms of risk-adjusted returns and long-term growth potential. It also compares both groups to the U.S. market to assess which region offers the most efficient contribution to a globally diversified portfolio.


# Setup and Data Access: Project Foundations

This section initializes the environment and defines functions to retrieve monthly time-series data from FRED and Alpha Vantage. These data sources provide the macroeconomic and asset-level information necessary to conduct portfolio return analysis and simulation.

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(kableExtra)
library(gganimate)
library(ggthemes)
library(httr)
library(jsonlite)
library(lubridate)
library(purrr)
library(PortfolioAnalytics)
library(quantmod)
library(PerformanceAnalytics)

#setwd("C:/Users/orazz/OneDrive - The City University of New York (1)/MASTER/STA9750-2025-SPRING")
```

- **Data wrangling**: `dplyr`, `tidyr`, `purrr`  
  Used to clean, manipulate, and reshape datasets efficiently.

- **Time handling**: `lubridate`  
  Simplifies working with dates in time series data.

- **Visualization**: `ggplot2`, `gganimate`, `ggthemes`, `kableExtra`.  
  Enables static and animated plotting, with themes to improve aesthetics.

- **API access**: `httr`, `jsonlite`  
  Used to send HTTP requests and parse JSON responses from the FRED and Alpha Vantage APIs.

- **Financial analysis and optimization**: `quantmod`, `PerformanceAnalytics`, `PortfolioAnalytics`  
  Provides tools for financial time series analysis, performance metrics, and portfolio optimization models.

## 🌐 Accessing FRED Data

The following code retrieves macroeconomic time series data (e.g., short-term interest rates or inflation) from the Federal Reserve’s FRED API.

- The function accepts a **FRED series ID**.

- It sends an HTTP request, parses the JSON response, and returns a clean `data.frame` with `date` and `value`.

- This is later used to calculate real returns or include macro indicators in portfolio modeling.

```{r}
FRED_key <- readLines("FRED_key.txt")
get_fred<- function(id){
  base_url <- "https://api.stlouisfed.org/fred/series/observations?series_id="
  res <- GET(paste0(base_url,id,"&api_key=",FRED_key,"&file_type=json"))
  res_content <- content(res, as = "text", encoding = "UTF-8")
  json <- fromJSON(res_content)
  data <-json$observations
  data <- data |> mutate(value = as.numeric(value),# immediately convert to usable format
                         date = as.Date(date))
  return(data)
}
```

## 📊 Alpha Vantage Data Function – Asset Prices

This function collects **monthly adjusted closing prices** from Alpha Vantage for a given ticker (e.g., SPY, IWM, GLD). 

- Returns a tidy data frame with adjusted close, dividend, and volume

- The adjusted close price accounts for splits and dividends, making it appropriate for return calculation.


```{r}
AV_key <- readLines("Alphavantage_key.txt")

# Function to get data from Alpha Vantage for a given ticker
GET_AV <- function(ticker){
  url <- paste0(
    "https://www.alphavantage.co/query?function=TIME_SERIES_MONTHLY_ADJUSTED",
    "&symbol=", ticker,
    "&apikey=", AV_key
  )
  
  res <- GET(url)
  res_content <- content(res, as = "text", encoding = "UTF-8")
  j <- fromJSON(res_content, flatten = TRUE)
  
  # Check if data exists
  if (is.null(j$`Monthly Adjusted Time Series`)) {
    print(j)  # to show what was returned
    stop(paste("No data returned for", ticker, "- possibly API limit reached or invalid ticker."))
  }
  
  data <- j$`Monthly Adjusted Time Series`
  
  df <- data.frame(
    date = as.Date(names(data)),
    close = as.numeric(sapply(data, `[[`, "4. close")),
    adjusted_close = as.numeric(sapply(data, `[[`, "5. adjusted close")),
    low = as.numeric(sapply(data, `[[`, "3. low")),
    volume = as.numeric(sapply(data, `[[`, "6. volume")),
    dividend = as.numeric(sapply(data, `[[`, "7. dividend amount"))
  )
  
  return(df)
}
```


## 📂 Data Caching Function

To avoid downloading the same data multiple times, I created a function that checks if the data for a specific ticker (like SPY or VXUS) is already saved on my computer.

- If it's not saved yet, it downloads the data and saves it as a CSV file.

- If it's already saved, it simply loads the data from the file.

- This helps save time and avoid hitting limits with the API. It also makes the analysis faster and more efficient.


```{r}
get_or_load_data <- function(ticker) {
  filename <- paste0(ticker, "_data.csv")
  
  if (!file.exists(filename)) {
    message("Fetching data for: ", ticker)
    data <- GET_AV(ticker)
    
    if (!is.null(data)) {
      write.csv(data, filename, row.names = FALSE)
    } else {
      warning("Data for ", ticker, " could not be retrieved.")
      return(NULL)
    }
  } else {
    message("Loading cached data for: ", ticker)
    data <- read.csv(filename) |>
      mutate(date = as.Date(date))  # ensure date column is in Date format
  }
  
  data$ticker <- ticker
  return(data)
}

# Now apply to each ticker
spy_data  <- get_or_load_data("SPY")
mdy_data  <- get_or_load_data("MDY")
iwm_data  <- get_or_load_data("IWM")
vxus_data <- get_or_load_data("VXUS")
vea_data  <- get_or_load_data("VEA")
vwo_data  <- get_or_load_data("VWO")
```

# 📈 Stock Market Data

I used the `get_or_load_data()` function to download or load historical price data for several key ETFs representing different segments of the global stock market:

- **SPY** – Tracks the S&P 500, representing large-cap U.S. stocks.

- **MDY** – Tracks the S&P MidCap 400, representing medium-sized U.S. companies.

- **IWM** – Tracks the Russell 2000, which includes small-cap U.S. stocks.

- **VXUS** – Covers the entire international stock market, excluding the U.S.

- **VEA** – Focuses on developed international markets (like Europe, Japan, and Canada).

- **VWO** – Focuses on emerging markets (like China, India, and Brazil).

This selection allows for comparison between U.S. equities and international equities, both developed and emerging, for analyzing long-term growth and diversification.

```{r}
# calculating log returns
spy_data <- spy_data |>
  arrange(date) |>
  mutate(log_return = log(adjusted_close / lag(adjusted_close)))

mdy_data <- mdy_data |>
  arrange(date) |>
  mutate(log_return = log(adjusted_close / lag(adjusted_close)))

iwm_data <- iwm_data |>
  arrange(date) |>
  mutate(log_return = log(adjusted_close / lag(adjusted_close)))

vxus_data <- vxus_data|>
  arrange(date)|>
  mutate(log_return = log(adjusted_close / lag(adjusted_close)))

vea_data <- vea_data|>
  arrange(date)|>
  mutate(log_return = log(adjusted_close / lag(adjusted_close)))

vwo_data <- vwo_data|>
  arrange(date)|>
  mutate(log_return = log(adjusted_close / lag(adjusted_close)))

combined_data_stock <- bind_rows(
  spy_data|> mutate(ticker = "SPY"),
  mdy_data|> mutate(ticker = "MDY"),
  iwm_data|> mutate(ticker = "IWM"),
  vxus_data|> mutate(ticker = "VXUS"),
  vea_data|> mutate(ticker = "VEA"),
  vwo_data|> mutate(ticker = "VWO")
)

```



# 🏦 Bond ETFs Used in the Analysis

I incorporated bond ETFs that track actual market prices and total returns. These ETFs offer a practical way to observe bond market performance using tradable instruments. Here's what each represents:

### 📘 BND – Vanguard Total Bond Market ETF

- **Description**: BND seeks to track the performance of the Bloomberg U.S. Aggregate Float Adjusted Bond Index.

- **Coverage**: A broad exposure to the entire U.S. investment-grade bond market, including U.S. Treasuries, agency bonds, mortgage-backed securities (MBS), and corporate bonds.

- **Use in Project**: Represents a diversified proxy for the U.S. bond market, useful for long-term risk and return comparison with equities and international assets.

### 📗 TLT – iShares 20+ Year Treasury Bond ETF

- **Description**: TLT tracks the performance of long-term U.S. Treasury bonds with maturities greater than 20 years.

- **Characteristics**: Highly sensitive to interest rate changes due to long duration; typically experiences more price volatility than short-term bonds.

- **Use in Project**: Provides a view into the performance of long-duration government debt, important for understanding interest rate risk.

### 📕 SHY – iShares 1–3 Year Treasury Bond ETF

- **Description**: SHY invests in U.S. Treasury bonds with maturities between 1 and 3 years.

- **Characteristics**: Lower duration and volatility; typically used as a safe, short-term investment vehicle.

- **Use in Project**: Serves as a low-risk benchmark to assess the risk-adjusted returns of longer-term bonds and equities.


```{r}

bnd_data <- get_or_load_data("BND")   # Total bond market
tlt_data <- get_or_load_data("TLT")   # Long-term Treasuries
shy_data <- get_or_load_data("SHY")   # Short-term Treasuries

# calculating log returns
bnd_data <- bnd_data |>
  arrange(date) |>
  mutate(log_return = log(adjusted_close / lag(adjusted_close)))

tlt_data <- tlt_data |>
  arrange(date) |>
  mutate(log_return = log(adjusted_close / lag(adjusted_close)))

shy_data <- shy_data |>
  arrange(date) |>
  mutate(log_return = log(adjusted_close / lag(adjusted_close)))

combined_bond_etfs<- 
  bind_rows(
    bnd_data|> mutate(ticker = "BND"),
    tlt_data|> mutate(ticker = "TLT"),
    shy_data|> mutate(ticker = "SHY")
  )
```




# 🏠 Commodities and Real Assets Data

To expand the asset classes in my portfolio analysis beyond just stocks and bonds, I included commodities and real estate data. These help diversify risk and potentially enhance returns across different market cycles. The ETFs I used are:

- **GLD**: Tracks the price of gold, commonly used as a hedge against inflation and market uncertainty.

- **DBA**: Represents a basket of agricultural commodities, such as wheat, corn, and soybeans, giving exposure to food production trends.

- **VNQ**: A Real Estate Investment Trust (REIT) ETF that provides exposure to U.S. real estate markets, offering income and growth potential separate from traditional equities.

```{r}
combined_data_commodities <- bind_rows(
  get_or_load_data("VNQ") |>
    arrange(date) |>
    mutate(ticker = "VNQ"),
  get_or_load_data("GLD") |>
    arrange(date) |>
    mutate(ticker = "GLD"),
  get_or_load_data("DBA") |>
    arrange(date) |>
    mutate(ticker = "DBA")
)

combined_data_commodities <- combined_data_commodities|> 
  group_by(ticker)|>
  mutate(log_return = log(adjusted_close / lag(adjusted_close)))|>
  ungroup()


combined_data <- bind_rows(
  combined_data_stock,
  combined_data_commodities,
  combined_bond_etfs
)


start_date <- as.Date("2011-01-01") 
combined_data <- combined_data |>
  filter(date >= start_date)

```



## 📈 Annualized Returns and Risk Metrics

To evaluate long-term performance across different asset classes—including U.S. stocks, international developed markets, and emerging markets—I calculated a set of standardized financial metrics using monthly log return data. The goal was to estimate both total and risk-adjusted returns over the observed period.

```{r}
yearly_returns <- combined_data |>
  drop_na(log_return) |>
  group_by(ticker) |>
  arrange(date) |>
  summarise(
    initial_price = first(adjusted_close),
    final_price   = last(adjusted_close),
    years         = as.numeric(difftime(last(date), first(date), units = "days")) / 365.25,
    total_return_pct = (final_price / initial_price - 1) * 100,
    CAGR_pct      = ((final_price / initial_price)^(1 / years) - 1)  ,
    volatility    = sd(log_return, na.rm = TRUE) * sqrt(12) ,
    sharpe_ratio  = (CAGR_pct  - 0.02) / (volatility )
  )

yearly_returns$CAGR_pct <- yearly_returns$CAGR_pct * 100
yearly_returns$volatility <- yearly_returns$volatility *100
```

### 🔍 Explanation of Code

- **`drop_na(log_return)`**: Removes missing values to ensure calculations are valid and consistent across all assets.
- **`group_by(ticker)`**: Performs all subsequent calculations separately for each asset (ETF) in the dataset.
- **`initial_price`, `final_price`**: Capture the first and last adjusted prices for each asset, forming the basis for return calculations.
- **`years`**: Calculates the total time span (in years) of the available data for each asset.
- **`total_return_pct`**: Computes the total percentage return over the entire period, from the initial to the final price.
- **`CAGR_pct`** *(Compound Annual Growth Rate)*: Reflects the mean annual growth rate of an investment over the time period, accounting for compounding.
- **`volatility`**: Measures the annualized standard deviation of the monthly log returns, serving as a proxy for investment risk.
- **`sharpe_ratio`**: Calculates the Sharpe Ratio, which adjusts the return by its risk. It uses a 2% annual risk-free rate to assess how efficiently an asset compensates investors for the risk taken.

These metrics collectively help compare the long-term performance and risk profile of U.S. markets, developed

## 📐 Financial Metric Formulas

### 1. Total Return (%)

$$
\text{Total Return \%} = \left( \frac{P_{\text{final}} - P_{\text{initial}}}{P_{\text{initial}}} \right) \times 100
$$


Where:  

- \( P_{\text{final}} \): Final price of the asset  
- \( P_{\text{initial}} \): Initial price of the asset

---

### 2. Compound Annual Growth Rate (CAGR %)

$$
\text{CAGR \%} = \left( \left( \frac{P_{\text{final}}}{P_{\text{initial}}} \right)^{\frac{1}{n}} - 1 \right) \times 100
$$

Where:  

- \( n \): Number of years the asset is held

---

### 3. Annualized Volatility

$$
\text{Volatility} = \sigma_{\text{monthly returns}} \times \sqrt{12}
$$

Where:

- \( \sigma_{\text{monthly returns}} \): Standard deviation of monthly log returns

---

### 4. Sharpe Ratio

$$
\text{Sharpe Ratio} = \frac{\text{CAGR} - r_f}{\text{Volatility}}
$$

Where:  
- \( r_f \): Risk-free rate (assumed to be 2% in this project)  
- Volatility: Annualized standard deviation of returns

# Assets Analysis{#assets-analysis}

# International Equity ETFs: VXUS vs. VEA vs. VWO

When building a diversified portfolio with international exposure, it's important to understand the different types of global equity ETFs. Below is a summary of three widely used ETFs and their key distinctions:

```{r ,echo=FALSE , output = TRUE}

yearly_returns|>filter(ticker %in% c("VXUS" , "VEA", "VWO"))|>
  select(-years)|> arrange(desc(CAGR_pct))|>
  rename("Initial Price" = initial_price,
         "Final Price" = final_price,
         "Total Return %" = total_return_pct,
         "CAGR %"=CAGR_pct,
         "Volatility" =volatility,
         "Ticker" = ticker,
         "Sharpe Ratio" = sharpe_ratio)|>
  kable(digits = 2, caption = "Comparison of International Market ETFs")
  
    
  
```

**VXUS – Vanguard Total International Stock ETF**

- Coverage: Broad exposure to all international markets outside the U.S., including both developed and emerging markets.

- Developed Markets: Includes countries like Japan, UK, Canada, and Europe.

- Emerging Markets: Includes countries like China, India, Brazil, and South Africa.

- Use Case: Ideal for investors seeking a single fund for global diversification beyond the U.S.

**VEA – Vanguard FTSE Developed Markets ETF**

- Coverage: Tracks developed international markets only, excluding the U.S. and Canada.

- Countries Included: Japan, UK, France, Germany, Australia, etc.

- No Emerging Markets Exposure

- Use Case: Suitable when the investor wants to separate developed and emerging market exposure.

**VWO – Vanguard FTSE Emerging Markets ETF**

-Coverage: Focused exclusively on emerging markets.

-Countries Included: China, India, Brazil, Taiwan, South Africa, etc.

-Higher Volatility, Higher Growth Potential

-Use Case: Used to target higher risk-adjusted returns and diversify geographically from developed markets.


```{r,fig.width=10, fig.height=7, output = TRUE}
rebased_data <- combined_data_stock|>
  filter(date >= as.Date("2011-01-01"))|>
  group_by(ticker) |> 
  arrange(date) |> 
  mutate(rebased = adjusted_close / first(adjusted_close) * 100)


rebased_data_int <- combined_data_stock|>
  filter(ticker %in% c("VXUS", "VEA", "VWO"))|> 
  filter(date >= as.Date("2011-01-01"))|>
  group_by(ticker) |> 
  arrange(date) |> 
  mutate(rebased = adjusted_close / first(adjusted_close) * 100)

rebased_data_int|>
  ggplot( aes(x = date, y = rebased, color = ticker)) +
  geom_line(linewidth = 1.) +
  labs(title = "Rebased International ETFs", 
       y = "Index (100 = Start)", 
       x = "Date",
       color = "ETF")+
  theme_bw()+
  scale_color_brewer(palette = "Set1")+
  scale_x_date(date_labels = "%Y", date_breaks = "3 year")+
  geom_hline(yintercept = 100, linetype = "dashed", color = "gray60")+
  #theme(legend.position = "bottom")+
  theme(text = element_text(family = "Helvetica", size = 15, face = "bold"))

  
```


```{r ,echo=FALSE , output = TRUE}

yearly_returns|>filter(ticker %in% c("SPY" ,"MDY","IWM", "VXUS" , "VEA", "VWO"))|>
  select(-years)|> arrange(desc(CAGR_pct))|>
  rename("Initial Price" = initial_price,
         "Final Price" = final_price,
         "Total Return %" = total_return_pct,
         "CAGR %"=CAGR_pct,
         "Volatility" =volatility,
         "Ticker" = ticker,
         "Sharpe Ratio" = sharpe_ratio)|>
  kable(digits = 2, caption = "US ETFs vs International Market ETFs")
  
```



```{r,fig.width=10, fig.height=7, output = TRUE}
rebased_data|> filter(ticker %in% c("SPY","MDY","IWM","VXUS" , "VEA", "VWO" ))|>
  ggplot( aes(x = date, y = rebased, color = ticker)) +
  geom_line(linewidth = 1.) +
  labs(title = "Rebased US vs International ETFs", 
       y = "Index (100 = Start)", 
       x = "Date",
       color = "ETF")+
  theme_bw()+
  scale_color_brewer(palette = "Set1")+
  scale_x_date(date_labels = "%Y", date_breaks = "3 year")+
  geom_hline(yintercept = 100, linetype = "dashed", color = "gray60")+
  #theme(legend.position = "bottom")+
  theme(text = element_text(family = "Helvetica", size = 15, face = "bold"))
  
```


# Bond Analysis 

```{r,echo=FALSE , output = TRUE}

yearly_returns|>filter(ticker %in% c("BND" , "TLT", "SHY"))|>
  select(-years)|>
  rename("Initial Price" = initial_price,
         "Final Price" = final_price,
         "Total Return %" = total_return_pct,
         "CAGR %"=CAGR_pct,
         "Volatility" =volatility,
         "Ticker" = ticker,
         "Sharpe Ratio" = sharpe_ratio)|>
  kable(digits = 2, caption = "Comparison of Bond ETFs")


```

```{r,fig.width=10, fig.height=7, output = TRUE}
rebased_data_bond <- combined_bond_etfs|>
  filter(date >= as.Date("2011-01-01"))|>
  group_by(ticker) |> 
  arrange(date) |> 
  mutate(rebased = adjusted_close / first(adjusted_close) * 100)

rebased_data_bond|> 
  ggplot( aes(x = date, y = rebased, color = ticker)) +
  geom_line(linewidth = 1.) +
  labs(title = "Rebased Bond ETFs", 
       y = "Index (100 = Start)", 
       x = "Date",
       color = "ETF")+
  theme_bw()+
  scale_color_brewer(palette = "Set1")+
  scale_x_date(date_labels = "%Y", date_breaks = "3 year")+
  geom_hline(yintercept = 100, linetype = "dashed", color = "gray60")+
  #theme(legend.position = "bottom")+
  theme(text = element_text(family = "Helvetica", size = 15, face = "bold"))


```

# Commodities Analysis 

```{r,echo=FALSE , output = TRUE}

yearly_returns|>filter(ticker %in% c("GLD" , "DBA", "VNQ"))|>
  select(-years)|>
  rename("Initial Price" = initial_price,
         "Final Price" = final_price,
         "Total Return %" = total_return_pct,
         "CAGR %"=CAGR_pct,
         "Volatility" =volatility,
         "Ticker" = ticker,
         "Sharpe Ratio" = sharpe_ratio)|>
  kable(digits = 2, caption = "Comparison of Commodities")


```

```{r,fig.width=10, fig.height=7, output = TRUE}

rebased_data_com <- combined_data_commodities|>
  filter(date >= as.Date("2011-01-01"))|>
  group_by(ticker) |> 
  arrange(date) |> 
  mutate(rebased = adjusted_close / first(adjusted_close) * 100)

rebased_data_com|> 
  ggplot( aes(x = date, y = rebased, color = ticker)) +
  geom_line(linewidth = 1.) +
  labs(title = "Rebased Commodities", 
       y = "Index (100 = Start)", 
       x = "Date",
       color = "ETF")+
  theme_bw()+
  scale_color_brewer(palette = "Set1")+
  scale_x_date(date_labels = "%Y", date_breaks = "3 year")+
  geom_hline(yintercept = 100, linetype = "dashed", color = "gray60")+
  #theme(legend.position = "bottom")+
  theme(text = element_text(family = "Helvetica", size = 15, face = "bold"))


```


```{r}
get_portfolio_weights <- function(age) {
  if (age < 30) {
    weights <- c(stocks = 0.90, bonds = 0.05, commodities = 0.05)
  } else if (age >= 30 & age <= 34) {
    weights <- c(stocks = 0.80, bonds = 0.10, commodities = 0.10)
  } else if (age >= 35 & age <= 39) {
    weights <- c(stocks = 0.75, bonds = 0.15, commodities = 0.10)
  } else if (age >= 40 & age <= 44) {
    weights <- c(stocks = 0.70, bonds = 0.20, commodities = 0.10)
  } else if (age >= 45 & age <= 49) {
    weights <- c(stocks = 0.65, bonds = 0.25, commodities = 0.10)
  } else if (age >= 50 & age <= 54) {
    weights <- c(stocks = 0.575, bonds = 0.30, commodities = 0.125)
  } else if (age >= 55 & age <= 59) {
    weights <- c(stocks = 0.50, bonds = 0.35, commodities = 0.15)
  } else if (age >= 60 & age <= 64) {
    weights <- c(stocks = 0.45, bonds = 0.40, commodities = 0.15)
  } else if (age >= 65 & age <= 69) {
    weights <- c(stocks = 0.40, bonds = 0.45, commodities = 0.15)
  } else {
    weights <- c(stocks = 0.35, bonds = 0.45, commodities = 0.20)
  }
  return(weights)
}

# Function to assign weights based on Sharpe ratios within each class
assign_weights_by_age <- function(yearly_returns, age) {
  allocation <- get_portfolio_weights(age)
  
  # Define ticker groups
  stock_tickers <- c("SPY", "MDY", "IWM", "VXUS", "VWO", "VEA")
  bond_tickers <- c("BND", "TLT", "SHY")
  commodity_tickers <- c("DBA", "VNQ", "GLD")
  
  # Tag each asset by class
  yearly_returns <- yearly_returns |>
    mutate(class = case_when(
      ticker %in% stock_tickers ~ "stocks",
      ticker %in% bond_tickers ~ "bonds",
      ticker %in% commodity_tickers ~ "commodities",
      TRUE ~ "other"
    ))
  
  # Separate bonds to handle fallback logic
  bonds <- yearly_returns |> filter(class == "bonds")
  other_assets <- yearly_returns |> filter(class != "bonds")
  
  # Check if all bond Sharpe ratios are <= 0
  all_bond_sharpe_negative <- all(bonds$sharpe_ratio <= 0 | is.na(bonds$sharpe_ratio))
  
  if (all_bond_sharpe_negative) {
    # Use yield-to-risk ratio for bonds if Sharpe ratios are all bad
    bonds <- bonds |>
      mutate(
        yield_to_risk = ifelse(volatility == 0, 0, CAGR_pct / volatility),
        raw_weight = yield_to_risk / sum(yield_to_risk, na.rm = TRUE),
        final_weight = raw_weight * allocation[["bonds"]]
      )
  } else {
    # Use Sharpe as usual
    bonds <- bonds |>
      mutate(
        adj_sharpe = ifelse(is.na(sharpe_ratio) | sharpe_ratio < 0, 0, sharpe_ratio),
        raw_weight = adj_sharpe / sum(adj_sharpe, na.rm = TRUE),
        final_weight = raw_weight * allocation[["bonds"]]
      )
  }
  
  # Handle other asset classes (stocks, commodities) with Sharpe
  other_assets <- other_assets |>
    filter(class %in% names(allocation)) |>
    group_by(class) |>
    mutate(
      adj_sharpe = ifelse(is.na(sharpe_ratio) | sharpe_ratio < 0, 0, sharpe_ratio),
      raw_weight = adj_sharpe / sum(adj_sharpe, na.rm = TRUE),
      final_weight = raw_weight * allocation[class]
    ) |>
    ungroup()
  
  # Combine and return
  bind_rows(bonds, other_assets) |>
    select(ticker, class, CAGR_pct, volatility, sharpe_ratio, final_weight)
}
```

# 📊 Adaptive Portfolio Allocation Based on Age

To make this portfolio model more realistic, I created a system that adjusts how much a person should invest in different asset classes (stocks, bonds, commodities) based on their age. Younger people can usually take more risk, while older individuals may want more stability.

### Set Recommended Allocations by Age

The function `get_portfolio_weights()` assigns a recommended percentage for each asset class depending on the user's age:

- **Younger investors (under 30)**: Heavily invested in stocks (90%), with small portions in bonds and commodities.

- **Older investors**: Gradually shift more money into bonds and commodities, which are considered safer. Going from **agressive** to **balanced** to **conservative** portfolio.  

This creates a simple life-cycle investing strategy that adjusts risk exposure over time.

### Adjust Portfolio Using Performance Metrics 

The `assign_weights_by_age()` function takes the age-based allocation and distributes it across individual assets (like SPY, TLT, GLD, etc.) using Sharpe ratios—a measure of risk-adjusted return.

- Assets with higher Sharpe ratios (better return per unit of risk) get more weight.

- If bond Sharpe ratios are too low or negative, the function switches to a backup method: using the CAGR-to-volatility ratio as a proxy for return-to-risk (called yield-to-risk).

This ensures that bad-performing bonds don’t drag down the portfolio just because of age-based rules.

### 🔧 How It Works
Group assets into: stocks, bonds, and commodities.

Filter by class, then assign weights:

- For bonds, use Sharpe ratios or yield-to-risk if Sharpe is poor.

- For stocks and commodities, use Sharpe ratios directly.

The final output is a table showing each ticker’s recommended weight in the portfolio based on the user's age and the asset’s historical performance.


```{r}

library(shiny)

# === UI ===
ui <- fluidPage(
  titlePanel("Age-Based Portfolio Allocation"),
  
  sidebarLayout(
    sidebarPanel(
      numericInput("age", "Enter your age (18–100):", value = 30, min = 18, max = 100),
      textOutput("error")
    ),
    
    mainPanel(
      tableOutput("weightsTable"),
      verbatimTextOutput("totalWeight")
    )
  )
)

# === SERVER ===
server <- function(input, output, session) {

  output$weightsTable <- renderTable({
    if (input$age < 18 || input$age > 100) {
      output$error <- renderText({ "Please enter an age between 18 and 100." })
      return(NULL)
    } else {
      output$error <- renderText({ "" })
    }

    # Use your actual yearly_returns data here
    result <- assign_weights_by_age(yearly_returns, input$age)

    # Add total row
    total_row <- result |>
      summarise(
        ticker = "TOTAL",
        class = "",
        CAGR_pct = NA,
        volatility = NA,
        sharpe_ratio = NA,
        final_weight = sum(final_weight, na.rm = TRUE)
      )

    bind_rows(result, total_row)
  })

  output$totalWeight <- renderPrint({
    if (input$age < 18 || input$age > 100) return()

    result <- assign_weights_by_age(yearly_returns, input$age)
    total_weight <- sum(result$final_weight, na.rm = TRUE)
    paste("Total portfolio weight:", round(total_weight, 4))
  })
}

# === RUN APP ===

shinyApp(ui = ui, server = server)


```





